---
title: "RawTools test subset Matrix output analysis"
output: html_notebook
---

This document details the analysis of Matrix data generated using RawTools for analysis of the subset of injections from the larger HeLa set (n = 10). 

<br>

### Data analysis and plotting

<br>

These are the R libraries we will require.

```{r, results = 'hide'}
library(ggplot2)
library(cowplot)
library(RColorBrewer)
library(dplyr)
library(data.table)
```

<br>

First we need to get the list of Matrix files that were output by RawTools.

```{r}
###################################################################################################################################################################
#get the list of metrics files
scan_files = list.files('/projects/ptx_analysis/chughes/projects-current/morin_rawtools/test_set/scanData', pattern = '.txt', full.names = TRUE)
```

<br>

Now we can process the data into a single data frame and prepare it for plotting.

```{r}
###################################################################################################################################################################
#read in the list of files and store in a single data frame
scan_all = data.frame()
for (i in 1:length(scan_files)){
  scan_in = read.table(scan_files[i], header = TRUE, sep = '\t')
  scan_in$fraction = sub(".*?Std_(.*?)\\.raw_Matrix\\.txt", "\\1", scan_files[i])
  scan_all = rbind(scan_all, scan_in)
}


###################################################################################################################################################################
#make fraction a factor
scan_all$fraction = factor(scan_all$fraction, levels = 1:10)
```

<br>

First plot a histogram of all the MS2 scan events.

```{r}
###################################################################################################################################################################
#plot a histogram of MS2 scans
frac3 = subset(scan_all, scan_all$fraction == 3)

output_plot = ggplot(frac3, aes(ddScanRetTime)) +
  geom_histogram(bins = 100, color = 'white', fill = brewer.pal(3,'Set3')[1]) +
  labs(x = "Retention Time (minutes)", y = 'Number of MS2 Scans', title = 'MS2 Scans Across Elution Window') +
  theme(axis.text.x = element_text(size = 9)) +
  scale_x_continuous(limits = c(0,30), breaks = seq(0,30,5))
output_plot
save_plot('/projects/ptx_analysis/chughes/projects-current/morin_rawtools/test_set/Ranalysis/fraction3_ms2-scan-numbers.pdf',output_plot)
```

<br>

Now look at MS1 intensities.

```{r}
###################################################################################################################################################################
#plot a line plot of MS1 intensities
frac3 = subset(scan_all, scan_all$fraction == 3)

output_plot = ggplot(frac3, aes(ddScanRetTime, MS1MedianIntensity)) +
  geom_line(color = brewer.pal(3,'Set3')[1]) +
  labs(x = "Retention Time (minutes)", y = 'Median MS1 Intensity', title = 'MS1 Intensity Profile') +
  theme(axis.text.x = element_text(size = 9)) +
  scale_x_continuous(limits = c(0,30), breaks = seq(0,30,5)) +
  scale_y_continuous(limits = c(0,6e5), breaks = seq(0,6e5,1e5))
output_plot
save_plot('/projects/ptx_analysis/chughes/projects-current/morin_rawtools/test_set/Ranalysis/fraction3_ms1-intensity.pdf',output_plot)
```

<br>

Number of MS2 scans per cycle, just for interest.

```{r}
###################################################################################################################################################################
#plot a temporal analysis of number of dependent scans
frac1 = subset(scan_all, scan_all$fraction == 1)

output_plot = ggplot(frac1, aes(ddScanRetTime, MS2ScansPerCycle)) +
  geom_point(color = 'black', fill = brewer.pal(3,'Set3')[1], pch = 21, size = 1, alpha = 0.6) +
  geom_smooth(method='loess',formula=y~x) +
  labs(x = "Retention Time (minutes)", y = 'Number of Dependent Scans', title = 'MS2 Scans Triggered per MS1') +
  theme(axis.text.x = element_text(size = 9)) +
  scale_x_continuous(limits = c(0,30), breaks = seq(0,30,5))
output_plot
save_plot('/projects/ptx_analysis/chughes/projects-current/morin_rawtools/test_set/Ranalysis/fraction1_dependent-scan-numbers.pdf',output_plot)
```

<br>

This last plot looks at the number of scans that a precursor was picked as a difference from the scan event where its max intensity was observed.

```{r}
###################################################################################################################################################################
#plot a temporal analysis number of scans away from the max intensity of a precursor
frac1 = subset(scan_all, scan_all$fraction == 1 & scan_all$ParentPeakFound == 'True')
frac1$scanDiff = abs(frac1$PeakMaxScan - frac1$MS1ScanNumber)

output_plot = ggplot(frac1, aes(ddScanRetTime, scanDiff)) +
  geom_point(color = 'black', fill = brewer.pal(3,'Set3')[1], pch = 21, size = 1, alpha = 0.6) +
  labs(x = "Retention Time (minutes)", y = 'Number of Dependent Scans', title = 'MS2 Scans Triggered per MS1') +
  theme(axis.text.x = element_text(size = 9)) +
  scale_x_continuous(limits = c(0,30), breaks = seq(0,30,5)) +
  scale_y_continuous(limits = c(0,50), breaks = seq(0,50,5))
output_plot
#save_plot('/projects/ptx_analysis/chughes/projects-current/morin_rawtools/test_set/Ranalysis/fraction1_dependent-scan-numbers.pdf',output_plot)
```

<br>

Done.




