---
title: "RawTools test subset Metrics output analysis"
output: html_notebook
---

This document details the analysis of Metrics data generated using RawTools for analysis of the subset of injections from the larger HeLa set (n = 10). 

<br>

### Data analysis and plotting

<br>

These are the R libraries we will require.

```{r, results = 'hide'}
library(ggplot2)
library(cowplot)
library(RColorBrewer)
library(dplyr)
library(data.table)
```

<br>

First we need to get the list of Metrics files that were output by RawTools.

```{r}
###################################################################################################################################################################
#get the list of metrics files
metrics_files = list.files('/projects/ptx_analysis/chughes/projects-current/morin_rawtools/test_set/metrics', pattern = '.txt', full.names = TRUE)
```

<br>

Now process the data for all 10 files into a single data frame and prepare for plotting.

```{r}
###################################################################################################################################################################
#read in the list of files and store in a single data frame
metrics_all = data.frame()
for (i in 1:length(metrics_files)){
  metrics_in = as.data.frame(t(read.table(metrics_files[i], header = FALSE, sep = '\t', row.names = 1)))
  metrics_in$fraction = sub(".*?Std_(.*?)\\.raw_Metrics\\.txt", "\\1", metrics_files[i])
  metrics_all = rbind(metrics_all, metrics_in)
}


###################################################################################################################################################################
#change the column names
colnames(metrics_all) = substring(names(metrics_all), 0, nchar(names(metrics_all))-1)
colnames(metrics_all)[24] = 'fraction'


###################################################################################################################################################################
#make fraction a factor
metrics_all$fraction = factor(metrics_all$fraction, levels = 1:10)
```

<br>

First plot the scan numbers across the replicates.

```{r}
###################################################################################################################################################################
#plot the scan numbers across replicates
metrics_all$TotalScans = as.numeric(as.character(metrics_all$TotalScans))
metrics_all$MS1Scans = as.numeric(as.character(metrics_all$MS1Scans))
metrics_all$MS2Scans = as.numeric(as.character(metrics_all$MS2Scans))

output_plot = ggplot(metrics_all) +
  geom_point(aes(fraction,TotalScans), colour = brewer.pal(3,'Set3')[1], size=4) +
  geom_point(aes(fraction,MS1Scans), colour = brewer.pal(3,'Set3')[3], size=4) +
  geom_point(aes(fraction,MS2Scans), colour = brewer.pal(4,'Set3')[4], size=4) +
  labs(x = "fraction", y = 'number of scans', title = 'scan numbers across replicates') +
  theme(axis.text.x = element_text(size = 9)) +
  geom_hline(yintercept=mean(metrics_all$TotalScans,na.rm=TRUE), color = brewer.pal(6,'RdBu')[1], alpha = 0.6) +
  geom_hline(yintercept=mean(metrics_all$MS1Scans,na.rm=TRUE), color = brewer.pal(6,'RdBu')[1], alpha = 0.6) +
  geom_hline(yintercept=mean(metrics_all$MS2Scans,na.rm=TRUE), color = brewer.pal(6,'RdBu')[1], alpha = 0.6) +
  scale_y_continuous(limits = c(0,12000), breaks = seq(0,12000,2000))
output_plot
save_plot('/projects/ptx_analysis/chughes/projects-current/morin_rawtools/test_set/Ranalysis/scan-number-across-replicates.pdf',output_plot)
```

<br>

Now plot the peak shape data across the replicates.

```{r}
###################################################################################################################################################################
#plot the peak shape metrics across the replicates
metrics_all$`MedianPeakWidthAt10Percent(s)` = as.numeric(as.character(metrics_all$`MedianPeakWidthAt10Percent(s)`))
metrics_all$`MedianPeakWidthAtHalfMax(s)` = as.numeric(as.character(metrics_all$`MedianPeakWidthAtHalfMax(s)`))

output_plot = ggplot(metrics_all) +
  geom_point(aes(fraction,`MedianPeakWidthAt10Percent(s)`), colour = brewer.pal(3,'Set3')[1], size=4) +
  geom_point(aes(fraction,`MedianPeakWidthAtHalfMax(s)`), colour = brewer.pal(3,'Set3')[3], size=4) +
  labs(x = "fraction", y = 'median peak width (s)', title = 'peak widths across replicates') +
  theme(axis.text.x = element_text(size = 9)) +
  geom_hline(yintercept=mean(metrics_all$`MedianPeakWidthAt10Percent(s)`,na.rm=TRUE), color = brewer.pal(6,'RdBu')[1], alpha = 0.6) +
  geom_hline(yintercept=mean(metrics_all$`MedianPeakWidthAtHalfMax(s)`,na.rm=TRUE), color = brewer.pal(6,'RdBu')[1], alpha = 0.6) +
  scale_y_continuous(limits = c(0,12), breaks = seq(0,12,1))
output_plot
save_plot('/projects/ptx_analysis/chughes/projects-current/morin_rawtools/test_set/Ranalysis/peak-width-across-replicates.pdf',output_plot)
```

<br>

Now the peak asymmetry.

```{r}
###################################################################################################################################################################
#plot the peak asymmetry across the replicates
metrics_all$MedianAsymmetryFactor = as.numeric(as.character(metrics_all$MedianAsymmetryFactor))

output_plot = ggplot(metrics_all) +
  geom_point(aes(fraction,MedianAsymmetryFactor), colour = brewer.pal(3,'Set3')[1], size=4) +
  labs(x = "fraction", y = 'median peak asymmetry factor', title = 'peak asymmetry across replicates') +
  theme(axis.text.x = element_text(size = 9)) +
  geom_hline(yintercept=mean(metrics_all$MedianAsymmetryFactor,na.rm=TRUE), color = brewer.pal(6,'RdBu')[1], alpha = 0.6) +
  scale_y_continuous(limits = c(0,2), breaks = seq(0,2,0.25))
output_plot
save_plot('/projects/ptx_analysis/chughes/projects-current/morin_rawtools/test_set/Ranalysis/peak-asymmetry-across-replicates.pdf',output_plot)
```

<br>

Peak capacity

```{r}
###################################################################################################################################################################
#plot the peak capacity across the replicates
metrics_all$ColumnCapacity = as.numeric(as.character(metrics_all$ColumnCapacity))

output_plot = ggplot(metrics_all) +
  geom_point(aes(fraction,ColumnCapacity), colour = brewer.pal(3,'Set3')[1], size=4) +
  labs(x = "fraction", y = 'column capacity', title = 'column peak capacity across replicates') +
  theme(axis.text.x = element_text(size = 9)) +
  geom_hline(yintercept=mean(metrics_all$ColumnCapacity,na.rm=TRUE), color = brewer.pal(6,'RdBu')[1], alpha = 0.6) +
  scale_y_continuous(limits = c(0,400), breaks = seq(0,400,50))
output_plot
save_plot('/projects/ptx_analysis/chughes/projects-current/morin_rawtools/test_set/Ranalysis/column-capacity-across-replicates.pdf',output_plot)
```

<br>

MS2 scan rate per second.

```{r}
###################################################################################################################################################################
#plot the MS2 Scan Rate per second
metrics_all$`MS2ScanRate(/sec)` = as.numeric(as.character(metrics_all$`MS2ScanRate(/sec)`))

output_plot = ggplot(metrics_all) +
  geom_point(aes(fraction,`MS2ScanRate(/sec)`), colour = brewer.pal(3,'Set3')[1], size=4) +
  labs(x = "fraction", y = 'MS2 scans per second', title = 'MS2 scan rate across replicates') +
  theme(axis.text.x = element_text(size = 9)) +
  geom_hline(yintercept=mean(metrics_all$`MS2ScanRate(/sec)`,na.rm=TRUE), color = brewer.pal(6,'RdBu')[1], alpha = 0.6) +
  scale_y_continuous(limits = c(0,8), breaks = seq(0,8,1))
output_plot
save_plot('/projects/ptx_analysis/chughes/projects-current/morin_rawtools/test_set/Ranalysis/MS2-scan-rate-across-replicates.pdf',output_plot)
```

<br>

Number of MS2 scans triggered per MS1.

```{r}
###################################################################################################################################################################
#plot the TopN Rate across replicates
metrics_all$MeanTopN = as.numeric(as.character(metrics_all$MeanTopN))

output_plot = ggplot(metrics_all) +
  geom_point(aes(fraction,MeanTopN), colour = brewer.pal(3,'Set3')[1], size=4) +
  labs(x = "fraction", y = 'mean TopN', title = 'TopN rate across replicates') +
  theme(axis.text.x = element_text(size = 9)) +
  geom_hline(yintercept=mean(metrics_all$MeanTopN,na.rm=TRUE), color = brewer.pal(6,'RdBu')[1], alpha = 0.6) +
  scale_y_continuous(limits = c(0,10), breaks = seq(0,10,1))
output_plot
save_plot('/projects/ptx_analysis/chughes/projects-current/morin_rawtools/test_set/Ranalysis/TopN-rate-across-replicates.pdf',output_plot)
```

<br>

Duty cycle duration.

```{r}
###################################################################################################################################################################
#plot the duty cycle across replicates
metrics_all$MeanDutyCycle = as.numeric(as.character(metrics_all$MeanDutyCycle))

output_plot = ggplot(metrics_all) +
  geom_point(aes(fraction,MeanDutyCycle), colour = brewer.pal(3,'Set3')[1], size=4) +
  labs(x = "fraction", y = 'mean duty cycle (s)', title = 'duty cycle across replicates') +
  theme(axis.text.x = element_text(size = 9)) +
  geom_hline(yintercept=mean(metrics_all$MeanDutyCycle,na.rm=TRUE), color = brewer.pal(6,'RdBu')[1], alpha = 0.6) +
  scale_y_continuous(limits = c(0,2), breaks = seq(0,2,0.25))
output_plot
save_plot('/projects/ptx_analysis/chughes/projects-current/morin_rawtools/test_set/Ranalysis/duty-cycle-across-replicates.pdf',output_plot)
```

<br>

Ion injection fill times for MS2 scans.

```{r}
###################################################################################################################################################################
#plot the fill times across replicates
metrics_all$MedianMS2FillTime = as.numeric(as.character(metrics_all$MedianMS2FillTime))

output_plot = ggplot(metrics_all) +
  geom_point(aes(fraction,MedianMS2FillTime), colour = brewer.pal(3,'Set3')[1], size=4) +
  labs(x = "fraction", y = 'median MS2 fill time (ms)', title = 'MS2 fill times across replicates') +
  theme(axis.text.x = element_text(size = 9)) +
  geom_hline(yintercept=mean(metrics_all$MedianMS2FillTime,na.rm=TRUE), color = brewer.pal(6,'RdBu')[1], alpha = 0.6) +
  scale_y_continuous(limits = c(0,10), breaks = seq(0,10,5))
output_plot
save_plot('/projects/ptx_analysis/chughes/projects-current/morin_rawtools/test_set/Ranalysis/fill-times-across-replicates.pdf',output_plot)
```

<br>

Lastly, MS2 intensities across the replicates. 

```{r}
###################################################################################################################################################################
#plot the MS2 intensities across replicates
metrics_all$MedianMS2Intensity = as.numeric(as.character(metrics_all$MedianMS2Intensity))

output_plot = ggplot(metrics_all) +
  geom_point(aes(fraction,MedianMS2Intensity), colour = brewer.pal(3,'Set3')[1], size=4) +
  labs(x = "fraction", y = 'median MS2 intensity', title = 'MS2 intensities across replicates') +
  theme(axis.text.x = element_text(size = 9)) +
  geom_hline(yintercept=mean(metrics_all$MedianMS2Intensity,na.rm=TRUE), color = brewer.pal(6,'RdBu')[1], alpha = 0.6) +
  scale_y_continuous(limits = c(0,3000), breaks = seq(0,3000,1000))
output_plot
save_plot('/projects/ptx_analysis/chughes/projects-current/morin_rawtools/test_set/Ranalysis/ms2-intensity-across-replicates.pdf',output_plot)
```

<br>

Done. 









