---
title: "RawTools and ProteoWizard MGF equivalency"
output: html_notebook
---

This document details the analysis of identification data generated using SearchGUI and PeptideShaker using RawTools and ProteoWizard generated MGF files. 

<br>

### Data analysis and plotting

<br>

These are the R libraries we will require.

```{r, results = 'hide'}
library(ggplot2)
library(cowplot)
library(RColorBrewer)
library(dplyr)
library(data.table)
```

<br>

First we need to get the output reports from PeptideShaker. Here we are using the Default PSM report export. 

```{r, results = 'hide'}
###################################################################################################################################################################
#get the list of psm files
rt_psmfiles = list.files('/projects/ptx_analysis/chughes/projects-current/morin_rawtools/test_set/rt_mgf', pattern = 'PSM_Report.txt', full.names = TRUE)
pw_psmfiles = list.files('/projects/ptx_analysis/chughes/projects-current/morin_rawtools/test_set/pw_mgf', pattern = 'PSM_Report.txt', full.names = TRUE)
```

<br>

Now we will process the identifications data to generate more manageable sets, as well as combining the RawTools and ProteoWizard data into a single object for downstream comparison.


```{r, results = 'hide'}
###################################################################################################################################################################
#read in the list of files and store in a single data frame
rt_set = data.frame()
for (i in 1:length(rt_psmfiles)){
  psm_temp1 = read.table(rt_psmfiles[i], header = TRUE, sep = '\t')
  psm_temp2 = psm_temp1 %>% 
    dplyr::select(Protein.s., Sequence, Spectrum.File, Spectrum.Title) %>% #select columns
    dplyr::mutate(Sequence, sequence = toupper(as.character(Sequence))) %>%
    dplyr::mutate(Protein.s., accession = sapply(strsplit(as.character(Protein.s.),","), '[', 1)) %>%
    dplyr::mutate(Spectrum.File, fraction = sub(".*?Std_(.*?)\\.raw\\.mgf", "\\1", Spectrum.File)) %>%
    dplyr::mutate(Spectrum.Title, scan = sub(".*?Spectrum_(.*?)", "\\1", Spectrum.Title)) %>%
    dplyr::select(accession, sequence, fraction, scan)
  rt_set = rbind(rt_set, psm_temp2)
}

pw_set = data.frame()
for (i in 1:length(rt_psmfiles)){
  psm_temp1 = read.table(pw_psmfiles[i], header = TRUE, sep = '\t')
  psm_temp2 = psm_temp1 %>% 
    dplyr::select(Protein.s., Sequence, Spectrum.File, Spectrum.Title) %>% #select columns
    dplyr::mutate(Sequence, sequence = toupper(as.character(Sequence))) %>%
    dplyr::mutate(Protein.s., accession = sapply(strsplit(as.character(Protein.s.),","), '[', 1)) %>%
    dplyr::mutate(Spectrum.File, fraction = sub(".*?Std_(.*?)\\.mgf", "\\1", Spectrum.File)) %>%
    dplyr::mutate(Spectrum.Title, scan = sapply(strsplit(as.character(Spectrum.Title),"\\."), '[', 2)) %>%
    dplyr::select(accession, sequence, fraction, scan)
  pw_set = rbind(pw_set, psm_temp2)
}

rt_set$dataset = 'rt'
pw_set$dataset = 'pw'
psm_set = rbind(rt_set, pw_set)
psm_set$dataset = factor(psm_set$dataset, levels = c('rt','pw'))
```

<br>

Now we can start doing some plotting. First let's look at PSM identification numbers. 

```{r}
###################################################################################################################################################################
#make a plot of the psm identifications
psm_set$psms = 1
psm_compile <- setDT(psm_set)[, lapply(.SD,sum,na.rm=TRUE),by=.(fraction, dataset),.SDcols=c('psms')]
setDF(psm_compile)

output_plot = ggplot(psm_compile, aes(dataset, psms)) +
  geom_boxplot(outlier.size = NA) +
  geom_jitter(size = 3, fill = 'grey60', alpha = 0.8, color = 'black', pch = 21) +
  labs(x = "Software Tool", y = 'Number of Peptides', title = 'Unique Peptide Matches') +
  theme(axis.text.x = element_text(size = 9)) +
  scale_y_continuous(limits = c(0,7000), breaks = seq(0,7000,1000))
output_plot
save_plot('/projects/ptx_analysis/chughes/projects-current/morin_rawtools/test_set/Ranalysis/mgf-validation_identifications-peptides.pdf',output_plot)

#t.test(psm_compile$psms~psm_compile$dataset, var.equal=TRUE, paired=FALSE)
```

<br>

What about protein identifications? 

```{r}
###################################################################################################################################################################
#make a plot of the protein identifications
pro_compile1 <- setDT(psm_set)[, lapply(.SD,sum,na.rm=TRUE),by=.(accession,fraction, dataset),.SDcols=c('psms')]
setDF(pro_compile1)
pro_compile1$proteins = 1
pro_compile2 <- setDT(pro_compile1)[, lapply(.SD,sum,na.rm=TRUE),by=.(fraction, dataset),.SDcols=c('proteins')]
setDF(pro_compile2)

output_plot = ggplot(pro_compile2, aes(dataset, proteins)) +
  geom_boxplot(outlier.size = NA) +
  geom_jitter(size = 3, fill = 'grey60', alpha = 0.8, color = 'black', pch = 21) +
  labs(x = "Software Tool", y = 'Number of Proteins', title = 'Unique Protein Matches') +
  theme(axis.text.x = element_text(size = 9)) +
  scale_y_continuous(limits = c(0,1300), breaks = seq(0,1300,200))
output_plot
save_plot('/projects/ptx_analysis/chughes/projects-current/morin_rawtools/test_set/Ranalysis/mgf-validation_identifications-proteins.pdf',output_plot)

#t.test(pro_compile2$proteins~pro_compile2$dataset, var.equal=TRUE, paired=FALSE)
```

<br>

Lastly, we can look at overlap in actual identifications between the two datasets for peptides and proteins.

```{r}
###################################################################################################################################################################
#compare protein overlap
psm_set$psms = 1
pro_compile <- setDT(psm_set)[, lapply(.SD,sum,na.rm=TRUE),by=.(accession, dataset),.SDcols=c('psms')]
setDF(pro_compile)

rt_proteins = subset(pro_compile, pro_compile$dataset == 'rt')
pw_proteins = subset(pro_compile, pro_compile$dataset == 'pw')

length(intersect(rt_proteins$accession, pw_proteins$accession))
length(union(rt_proteins$accession, pw_proteins$accession))

###################################################################################################################################################################
#compare peptide overlap
psm_set$psms = 1
pep_compile <- setDT(psm_set)[, lapply(.SD,sum,na.rm=TRUE),by=.(sequence, dataset),.SDcols=c('psms')]
setDF(pep_compile)

rt_peptides = subset(pep_compile, pep_compile$dataset == 'rt')
pw_peptides = subset(pep_compile, pep_compile$dataset == 'pw')

length(intersect(rt_peptides$sequence, pw_peptides$sequence))
length(union(rt_peptides$sequence, pw_peptides$sequence))
```

<br>

Done.





